---
title: "Getting Started with omicslog"
output:
  BiocStyle::html_document:
    toc: true
    toc_float: true
vignette: >
  %\VignetteIndexEntry{Getting Started with omicslog}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = TRUE,
  warning = TRUE
)
```

# Installation

You can install the `omicslog` package from Bioconductor using the following commands:

```{r, eval=FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install("omicslog")
```

# Introduction

```{r setup, message=FALSE}
library(SummarizedExperiment)
library(tidySummarizedExperiment)
library(omicslog)
data(airway, package="airway")
```

The `omicslog` package provides logging capabilities for `SingleCellExperiment` and `SummarizedExperiment` objects. 
This is particularly useful for tracking transformations in complex analysis workflows.

## Pipeline Workflow

Alternatively, you can chain these operations into a single pipeline:

```{r, message=TRUE, warning=TRUE}
# Complete workflow in a single pipeline
result <- 
  airway |>
  log_start() |>
  filter(dex == "untrt") |>
  select(!albut) |>
  mutate(dex_upper = toupper(dex)) |>
  extract(col = dex,into = "treat") |>
  mutate(Run = tolower(Run)) |> 
  filter(.feature == "ENSG00000000003") |>
  slice(3)


# View the object with the top log history as text
result
```

```{r}
# View the table with the complete log history
print(result@log_history,width=Inf)
```

## Base R Pipeline

Here's the same workflow implemented using base R operations:

```{r, message=TRUE, warning=TRUE}
# revert printing to Bioconductor style
options(restore_SummarizedExperiment_show = TRUE)

# Start with logging
result_base <- log_start(airway)

# Filter samples by dex
result_base <- result_base[, colData(result_base)$dex == "untrt"]

# Add new column with uppercase dex
colData(result_base)$dex_upper <- toupper(colData(result_base)$dex)

# Modify Run column to lowercase
colData(result_base)$Run <- tolower(colData(result_base)$Run)

# Filter features
result_base <- result_base[rownames(result_base) == "ENSG00000000003", ]

# View the object with the top log history as text
result_base
```

```{r}
# View the table with the complete log history
print(result_base@log_history,width=Inf)
```

```{r echo=FALSE}
# revert printing to Bioconductor style
options(restore_SummarizedExperiment_show = FALSE)
```

# SingleCellExperiment Example

`omicslog` also supports `SingleCellExperiment` objects. Here's an example using the `pbmc_small` dataset:

```{r, message=TRUE, warning=TRUE}
data(pbmc_small, package="tidySingleCellExperiment")
pbmc_logged <- log_start(pbmc_small)
pbmc_logged <- pbmc_logged[, colData(pbmc_logged)$groups == "g1"]
pbmc_logged <- pbmc_logged[rownames(pbmc_logged) %in% c("ENSG00000198888", "ENSG00000198786"), ]
pbmc_logged
```

# Session Info

```{r}
sessionInfo()
``` 