---
title: "Getting Started with omiclog"
author:
  - name: Your Name
    affiliation: Your Institution
    email: your.email@example.com
output: BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{Getting Started with omiclog}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

```{r setup}
library(omiclog)
library(SummarizedExperiment)
library(tidySummarizedExperiment)
```

The `omiclog` package provides logging capabilities for `SummarizedExperiment` objects. 
This is particularly useful for tracking transformations in complex analysis workflows.

# Basic Usage

## Creating a Logged SummarizedExperiment

To start logging, simply apply the `log()` function to any `SummarizedExperiment` object:

```{r}
# Load example dataset
data <- tidySummarizedExperiment::pasilla

# Start logging
logged_data <- log(data)
```

## Filtering Operations

When you filter a logged object, information about the operation is automatically tracked:

```{r}
# Filter to keep only treated samples
treated <- logged_data |>
  filter(condition == "treated")

# Filter to a specific gene
gene_of_interest <- treated |>
  filter(.feature == "FBgn0000003")
```

The log shows how many samples or genes were removed and what percentage they represent.

## Column Modifications

Column additions and modifications are also tracked:

```{r}
# Add a log-transformed counts column
transformed <- gene_of_interest |>
  mutate(log_counts = log2(counts + 1))
```

The log shows which columns were added or modified.

## Viewing the Log

The log is automatically displayed when the object is printed:

```{r}
# View the object with its log
transformed
```

# Advanced Usage

## Working with the Log History

The log history is stored in the object and can be accessed directly:

```{r}
# Access the log history
log_history <- transformed@log_history
print(log_history)
```

This allows you to programmatically work with the log if needed.

# Session Info

```{r}
sessionInfo()
``` 